---
title: "Data Notes"
output:
  html_document:
    collapsed: yes
    number_sections: no
    smooth_scrol: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
  )
library(mscatch)
```

# ADIOS!

This is the database that the population dynamics (Pop Dy) branch use for their stock assessments.  

* Access the online [GUI](file://155.206.139.2/home0/pdy/pub/STOCKEFF/ADIOS/website/webfiles/ADIOS_homepage.html)

* Read the [DOCUMENTATION](file://155.206.139.2/home0/pdy/pub/STOCKEFF/ADIOS/ADIOS_CF/website/webfiles/ADIOS_technical_documentation.pdf)

The landing portion are stored in the Stockeff.MV_cf_Landings materialized view. (This is equivalent to AA Trip and AA Species tables joined). Each record represents trip and species data for a specific landed species. A trip will appear as many times as there are landed species in that trip

* STOCKEFF.MV_CF_LANDINGS  - (WOLANDT/S,WODET/S, CFDETT/S tables combined)

* STOCKEFF.MV_CF_AGE - (WOAGE, CFAGE AA tables combined)

* STOCKEFF.MV_CF_LEN (WOLEN, CFLEN AA tables combined)

* CFDBS.CFSPP (MKTCAT, MKTNM)

## Statistical Area

In the database there are:

* 35,515,060 records
* 6,913,365 record (20%) still have unknown area designation (AREA = '000')

Pop Dy branch will assign these records to areas based on the species of interest. Of course this means that a trip could be assigned to several different areas depending on the species of interest. The algorithm for assigning the area is run "behind the scenes" when a scientist using the web GUI makes a request for data. [Methodology](file://155.206.139.2/home0/pdy/pub/STOCKEFF/documentation/references/Palmer_2008_unknown_apportionment.pdf) by Mike Palmer


## Latitude and Longitude

* 35,515,060 records
* 14,288,139 records (40%) have valid GIS_LAT, GIS_LON
* 21,227,636 records (60%) where GIS_LAT, GIS_LON is null

* 1,952,157 records (5%) where LAT, LON are null 
* 1,482,958 records (4%) where LAT or LON are '00'
* 17,791,806 records (50%) where LAT or LON are '0000'

### Trip level records

Use link field as the unique identifier

* 15,774,961 unique trips (15,776,348 in AA tables -> 1387 trips removed)
* 2,596,744 records (16%) have valid GIS_LAT, GIS_LON

There is a TENMSQ field but only included from 1994+


Allocation rules:

1. All trips WITH missing AREA (='000') need to be assigned to an AREA

1. If most of the GIS_LAT and GIS_LON fields contain valid entires we can allocate these TRIPS to an EPU.

1. For those that do not (84%) we will need to allocate these trips to an EPU based on the AREA they fished.

1. If they fished in an AREA that is entirely within an EPU, we can allocate to that EPU

1. Otherise the AREA straddles multiple EPUs so we will need to look at all trips already in associated with that AREA and allocate to an EPU based on most similar trips. Proceed in a similar fashion to how Sean Lucey allocates trips to missing AREA (step 1)


## Market Category

Simplified to categories based on the species (Looking for definitions)

eg Cod -> "UN", "LG", "MK", "SK"

All market category descriptions are unique to each species. 
eg. 
Mackerel (NESPP3='212')
```
channel = cfdbs::connect_to_database("sole",username)
table <- RODBC::sqlQuery(channel,"select nespp4,mktnm from cfdbs.CFSPP where nespp3 = '212' order by nespp4")
```
vs Cod (NESPP3 = '081')
```
channel = cfdbs::connect_to_database("sole",username)
table <- RODBC::sqlQuery(channel,"select nespp4,mktnm from cfdbs.CFSPP where nespp3 = '081' order by nespp4")
```



## Species (species_itis) {.tabset .tabset-pills}

Insignificant amount of missing species info (NULL, '000' or, '999') 
Most assessments start in either 1982 (rec fishing data starts) or 1989 (discard data starts). Often discards are back calculated in some way from 1982-1988 when needed.

``` {r specieslookup, echo=F,eval=T}
knitr::kable(speciesLookupTable)
```

### Spiny Dogfish

Kathy Sosebee

nespp3 = 352  (cfdbs)
svspp =  (svdbs)
species_itis = 160617

dogfish(nk) - 350. Are these used in the assesment

Mean length over all samples is calculated and applied to landings. No length expansion done

### winter skate

Kathy Sosebee

nespp3 = 367 (365 too? Rajidae)
species_itis = 564145 (160845 Rajidae)

Multiple market categories. Large (for wings), small for something else (size cutoff about 60cm)
Skate complex. Apportion out different species based on Survey landings

### Herring

John Deroba

nespp3 = 168, species_itis = 161722

[SAW 54 CRD12-18 (2012)](https://www.nefsc.noaa.gov/publications/crd/crd1218/) p48 explains how data were collected and aggregated

### Cod

Mike Palmer

nespp3 = 081 (082 ALSO? 400+ rows) 
species_itis = 164712

### Pollock

Brian Linton


### Haddock 

According to G. Shepherd has most complete data (landings, length, discards).
This will be trial dataset.

Meet with Liz Brooks week of 15th april 2019

nespp3 = 147 (148 too? Market Codes X1,X2,MK,MD, 0 length samples small landings. Appear in later years 2002+)
species_itis = 164744

#### length-weight (SAW 2014 ,p 31)
Length in cm, weight in kg

 $$ Wlive = 0.000007690·L_{fork} ^ {3.0622} $$ (spring)
 
 $$ Wlive = 0.000009870·L_{fork} ^ {3.0090} $$ (fall)
 
 $$ Wlive = 0.000009298·L_{fork} ^ {3.0205} $$ (annual)
 

### Yellowtail flounder

Chris Legault (GB)
Assessment consists of Biomass from survey adjusted by catchability (q = .31). Then an exploitation rate applied to inform catch advice. No model is used

nespp3 = 123
species_itis = 172909

### Winter flounder

Paul Nitschke (GOM)
Lisa Hendrickson (GB)

nespp3 = 119, (120 - 0 records)
species_itis = 172905

### Mackerel

K. Curti assessment from 1989.

Result - aggregated to 2 areas (NE, MA) by semester. Really only 1 gear type

nespp3 = 212
species_itis = 172414

Data may come from other sources also. NAFO.

Prior to 1992 Joint venture. CF tables landings dont match up with prior assessments since Joint venture data is stored in databases (servers = FishA, FishB). (Tables something like FFJV) which are old and were not converted to oracle databases. So to tease them apart from US landings is near impossible. For mackerel assessment data from < 1992 found in prior assessments are assumed as true since recreation is not possible.

### White Hake

Kathy sosebee

Length expansions to YEAR, SEMESTER, MARKET_CODE. Same structure for all years (many size classes icluding headed and gutted)

### silver hake

Larry Alade: Hake species

nespp3 = 509
species_itis = 164791

### Red Hake

Toni Scute (Kathy Sosebee past stock assessment scientist).

nespp3 = 
species_itis = 

(Red hake scoping group:11/04/2019)
Index based assessement (AIM). Currently in scoping phase to identify stock structure for a 2019 assessment. Current management unit describes a northern and southern stock. Boundary right through middle of Georges Bank. Prior to 1986 there were 4 stock areas. 2010 was last time stock stucture was examined. Since then new data available. (fisheries dependent (observer, VTR) data, survey (NEFSC Trawl, GOM shrimp, NEAMAP, NEFSC scallop, GOM longline (New data), canadian), life history). 

Working on catchability estimates for rock hopper sweep (twin sweep)

### gooosefish

Ann Richards

nespp3 = 011, 012
species_itis = 164499


### Market Code Lookup Table

``` {r itis}
knitr::kable(marketCodeLookupTable)

```

## Gear

How deal with missing gear info? How much is there ? 

## Time

Some months = 00 but have landings data. what do with them?


# Questions for Stock Assessment Scientists

1. For areas defined as '000' (post AA table creation). How do they deal with them?
For example: Does the scientist pull all cod data then deal with the missing areas from this reduced data set or are all missing areas dealt with THEN cod data are pulled?

1. Is this consistent for all scientists? and is the method of dealing with missing area  designations consistent?

1. If everyone is using ADIOS then what are the scripts in the background (from the GUI) doing to answer these questions?

1. 
