---
title: "Data Requirements"
#output: html_document
output: rmarkdown::html_vignette
csl: apa-annotated-bibliography.csl
bibliography: references.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{Data Requirements}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
 library(magrittr)
```

To use `mscatch` you will need to compile two data sets for a specific region of interest, whether a stock area, an ecological production unit, or a larger area:

* Length data
* Catch data

If you want a sample data set to begin to use `mscatch` then you can use the function [`get_sample_data`](../reference/get_sample_data.html)

## Length data

Length data (lengths of sampled catch taken by port samplers) can be pulled from the oracle table (`stockeff.mv_cf_len`) under the stock efficiency initiative or from the annual (cflen, wolen) AA tables. It is then aggregated by YEAR, QTR, GEAR, MARKET_CODE to summarize two data sets

* The size of the biological samples taken on trips
* Individual fish lengths and their frequencies from these samples

### Size of biological samples

* The number of length samples (number trips in which samples were taken) 
* The total number of lengths (number of actual fish that were sampled) 

This is then merged with the catch data.

### Individual fish lengths

* The individual lengths of a sampled fish (1 cm increments)
* The number of times a fish of this length was sampled

The data set should have this form

``` {r lengths, echo = F, eval = T}
sl <- readRDS(here::here("data-raw/data/sampleLengths.RDS")) %>%
  dplyr::select(-tripid)
sl
```

where each row represents the number of fish sampled of a given length in YEAR, QTR, caught using NEGEAR and classed as MARKET_CODE

* LENGTH is the length in cm of sampled fish and 
* NUMLEN is the number of fish of this size sampled 

## Catch data

Landings data can be pulled from the oracle table (`stockeff.mv_cf_landings`) under the stock efficiency initiative or from the annual (cfdet, wodet, woland) AA tables. If pulled from the AA tables, the MARKET_CODE field does not exist. NESPP4 codes will need to be mapped to MARKET_CODE categories outlined `mscatch::marketCodeLookupTable`. (created from [`comlandr::get_species_itis`](https://noaa-edab.github.io/comlandr/reference/get_species_itis.html))

Some trips will have missing information in the AREA field. These trips will need to be assigned an area using the methods outlined by [@Palmer2008] prior to using `mscatch`. The package [`comlandr`](https://noaa-edab.github.io/comlandr/index.html) replicates these methods.

Trips with missing gear information is handled in `mscatch`. It is assumed that trips with unknown gear information are catching fish with lengths proportional to trips with known gear types.

Discards are not included in this data. At a future date both [`comlandr`](https://noaa-edab.github.io/comlandr/index.html) and the stock efficiency initiative will contain discard estimates using the methods of [@Wigley2020].

The resulting Catch data is then merged with the biological samples to summarize total catch (metric tons) by YEAR, QTR, GEAR, MARKET_CODE, the number of biological samples taken (`landings_nn`), and the total number of fish sampled (`len_totalNumlen`). The data set should have this form:

``` {r catch, echo = F, eval = T}
sc <- readRDS(here::here("data-raw/data/sampleCatch.RDS"))
sc
```


## Use `mscatch`

Once the above data sets have been assembled they can now be used to [aggregate the landings and length samples](2aggregationRules.html) and then [expand](3catchExpansion.html) the commercial landings data to length distributions using `mscatch`


## References

