---
title: "mscatch"
#output: html_document
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mscatch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
 library(magrittr)
source(here::here("data-raw/R/create_species_rules_table.r"))
```

When working with `mscatch`, depending on what you want to do, you will need to:

*   [Assemble data](../articles/dataNeeds.html) for use in the package
*   [Aggregate the catch data](../articles/aggregationRules.html) based on a set of rules
*   Use length-weight relationships to [expand the catch](../articles/catchExpansion.html) to weights using the length samples
*   [Create an Age-Length Key](../articles/ageExpansion.html)
*   Calculate [numbers at age](articles/numbersAtAge.html)


Assuming you have assembled the data (for a sample data set see,  [`get_sample_data`](../reference/get_sample_data.html)), the next step is to determine how the data should be aggregate and how to deal with missing length samples in YEARs/QTRs where there are catch.

There are two approaches a user can take:

* Define specific rules in a species object list
* Specify general rules for a more exploratory approach.

### Define a species object

A [species object](https://noaa-edab.github.io/mscatch/reference/get_species_object.html) can be defined and passed as an argument to the [aggregate_landings](https://noaa-edab.github.io/mscatch/reference/aggregate_landings.html) function. This object specifies decision rules to be used in the aggregation step that answer the following questions:

* Which gear types should be aggregated?
* Which market codes should be aggregated?
* How to deal with with missing length samples? where to borrow from if at all?
* Which statistical areas define the stock?
* Where is length-weight data pulled from to fit length-weight relationships? survey database, commercial database, both?
* At what time scale are length weight relationships fitted? Semester, quartery, annually?
* Where is the age data pulled from?
* How should age length keys be calculated? Annually, by semester annually, by gear, semester, annually?
* What is the maximum age of the species ?

### General rules

This option is for exploring the data and assessing the impacts related to aggregation decisions to help inform building a species object. For example:

* To explore the most important gear types that are associated with landing a species a user can specify a threshold value relating to the proportion of cumulative landings, over which all other gear types are considered insignificant and are grouped together.

* Market codes are aggregated if the length frequency distributions are found to be significantly insignificant (using the Kolmogorov-Smirnov test). The level of the test (p-value) can be passed as an argument.


## Examples

### Using the exploratory approach


* Assigns the name of species (`speciesName = "Haddock"`)

* Aggregates landings over QTR year to YEARly data and combines all length samples to same level of aggregation (`aggregate_to = "YEAR"`).

* Uses gear types (NEGEARs) which contribute the most to total landings. Uses a cumulative landings threshold of 90% to select gears. Other gears will be aggregated into an "other Gear" category (`landingThresholdGear = 0.9`)

* Assigns other gears to code "998" (`otherGear = "998"`)

* Borrows length samples from nearest neighbor, whether by QTR, YEAR, or NEGEAR (`borrowLengths = T`). 

* Minimum number of length samples required when borrowing samples from other NEGEARs, QTRs, or YEARs (`nLengthSamples = 1`)

* Assigns an output folder (`outputDir=here::here("output/Haddock")`) for all diagnostic plots and the logfile

* Assigns a name to the log file produced (`logfile = "logfile.txt"`). This documents all decisions made in aggregation steps and documents the years in which length samples were borrowed and the year from which they were borrowed


```{r aggdata, echo =T, eval = F}
 aggregatedData <- mscatch::aggregate_landings(channel,
                                               landingsData,
                                               lengthData,
                                               outputDir=here::here("output/Haddok"),
                                               speciesName = "Haddock"
                                               landingThresholdGear = 0.9,
                                               nLengthSamples = 1,
                                               aggregate_to = "YEAR",
                                               borrowLengths = T,
                                               otherGear = "998",
                                               logfile = "logfile.txt"
                                               outputPlots = T,
                                               speciesRules = NULL)
```

At this point, `aggregateData` contains length samples for every combination of YEAR, NEGEAR, MARKET_CODE. All decisions made regarding gear type selection, market code selection, and which YEAR length samples were borrowed from are recorded in the `logfile`. If `outputPlots=T` a suite of plots are also output in the `outputDir`.

### Using a species object

A template for the species object is bundled with `mscatch` which can then be adapted

```{r template , echo = T, eval = T}
# Haddock species itis code = 164744
species_itis <- 164744
speciesRules <- mscatch::get_species_object(species_itis)
speciesRules$marketCodes %>% 
  flextable::flextable()
```


This specifies the market code aggregation; LG and XG codes are combind and relabeled as LG, SR and SK remain as they are. All other market codes are treated as unclassifieds, UN.

```{r gr}
speciesRules$gearCodes %>% 
  flextable::flextable()
```

All gear codes are aggregated and labelled as gear code "050"


Other related rules:

```{r otherprops, echo = F, eval = T}
create_species_rules_table(spitis = species_itis)

```

This `speciesRules` object is then passed as an argument to  [`aggregate_landings`](../reference/aggregate_landings.html)


### Expand the catch

To expand the catch to weight-by-length a length-weight key is required. This is achieved by obtaining parameter values that satisfy the [length-weight relationship](catchExpansion.html). This can be obtained in one of two ways: 

* Pull parameter values from Survey database (svdbs) using function [`get_length_weight`](https://noaa-edab.github.io/survdat/reference/get_length_weight.html) function from the [`survdat`](https://noaa-edab.github.io/survdat/index.html) package.

or

* Pull raw length-weight data and fit own model. The bottom trawl survey data is used (for this example). A series of nested length-weight models are then fit to the data.

The data is pulled using the function [`get_length_weight_data`](https://noaa-edab.github.io/survdat/reference/get_length_weight_data.html) from the [`survdat`](https://noaa-edab.github.io/survdat/) package and fit using `mscatch::fit_length_weight()`

```{r lw, echo =T , eval = F}
# Permissions and credentials are required to conect to the database
channel <- dbutils::connect_to_database("server","user")
# Haddock data are pulled (SVSPP = 74)
lengthWeightData <- survdat::get_length_weight_data(channel,year="all", species=74)
# Length-Weight models are thenn fit
fits <- mscatch::fit_length_weight(lengthWeightData$data,speciesName,speciesRules,outputDir,logfile="logfile.txt")

```

* The fitted model is then used to expand the catch. This is achieved by using `mscatch::expand_landings_to_lengths()`

```{r expand, echo = T, eval = F}
  expandedLandings <- mscatch::expand_landings_to_lengths(aggregatedData$landings,aggregatedData$lengthData,fits$paramsH0) %>%
    dplyr::mutate(speciesName = speciesName)

```


* Pull age data from either survey, commercial or both

* Create age-length key

* Calculate the numbers at age
