---
title: "mscatch"
#output: html_document
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mscatch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
 library(magrittr)
```

When working with `mscatch` you will need to:

*   [Assemble data]() for use in the package
*   [Aggregate the data]() based on a set of rules
*   Use length-weight relationships to [expand the lengths]() to weights

## Example

Assuming you have data assembled the first step is determine how you would like to aggregate the data and how you would like to deal with YEARs/ QTRs with missing samples. 

The following example implements the following decisions:

* Assign the name of species (`speciesName = "Haddock"`)

* Aggregate landings over QTR year to YEARly data and combine all length samples to same level of aggregation (`aggregate_to = "YEAR"`).

* Only use gear types (NEGEARs) which contribute the most to total landings. Use a cumulative landings threshold of 90% to select gears. Other gears will be aggregated into an "other Gear" category (`landingThresholdGear = 0.9`)

* Assign other gears to code "998" (`otherGear = "998"`)

* Borrow length samples from nearest neighbor, whether by QTR, YEAR, or NEGEAR (`borrowLengths = T`). 

* Minimum number of length samples required when borrowing samples from other NEGEARs, QTRs, or YEARs (`nLengthSamples = 1`)

* Assign an output folder (`outputDir=here::here("output/Groundfish")`)

* Assign a name to the log file produced (`logfile = "logfile.txt"`)


```{r , echo =T, eval = F}
 aggregatedData <- mscatch::aggregate_landings(landingsData,
                                               lengthData,
                                               outputDir=here::here("output/Groundfish"),
                                               speciesName = "Haddock"
                                               landingThresholdGear = 0.9,
                                               nLengthSamples = 1,
                                               aggregate_to = "YEAR",
                                               borrowLengths = T,
                                               otherGear = "998",
                                               logfile = "logfile.txt"
                                               outputPlots = T)
```

At this point for every combination of YEAR, NEGEAR, MARKET_CODE in `aggregateData` length samples exist. This is the requirement to expand the catch data to length composition.

* Obtain length-weight parameters that satisfy the [length-weight relationship](). For this example we will use the bottom trawl survey to pull lengths and weights for individual fish then use the data to fit the model.

The data is pulled using the [`survdat`](https://noaa-edab.github.io/survdat/) package and fit using `mscatch::fit_length_weight()`

```{r lw, echo =T , eval = F}
# You will need permissions and credentials to conect to the database
channel <- dbutils::connect_to_database("server","user")
# Pull the data for Haddock
lengthWeightData <- survdat::get_length_weight_data(channel,year="all", species=74)
# fit a model to the data
fits <- mscatch::fit_length_weight(lengthWeightData$data,speciesName,outputDir,logfile="logfile.txt")

```

* The fitted model is then used to expand the catch. Ths is achieved by using `mscatch::expand_landings_to_lengths()`

```{r expand, echo = T, eval = F}

```


